### 2019-08-26 chapter_02_open_science.tex
Redo figures and a final proofread

Collaboration was achieved in the early years of FOSS development through
emailing ``patches'', and centralized Version Control Systems (VCS) near the
turn of the 21\textsuperscript{st} century.  The collective collaboration on
development is today streamlined with the emergence of Distributed Version
Control Systems (DVCS), especially, Git and Mercurial).
%
Version control tools are designed to work on any kind of text files, not just
codes, so they could even be utilized to write scientific documents with
multiple authors.
%
In web-based source development platforms like GitHub, Bitbucket and GitLab,
DVCS repositories are enhanced with code-review tools, typically consisting of
a issue tracker and pull-request mechanism.
%
Git has become the \textit{de facto} standard for source management. However,
for the FluidDyn project, we prefer to use Mercurial because we consider it to
be simpler for the beginners and just as powerful for expert users\footnote{See
  \href{https://www.mercurial-scm.org/wiki/GitConcepts}{``a detailed
comparison''} on Mercurial website.}. Simplicity and low learning barriers for
scientists who are not experts in software development is crucial for our
project.
%
Note that, it is now possible to use Mercurial to work with Git repositories,
for example hosted in GitHub and GitLab\footnote{By using the Mercurial
extension \href{http://hg-git.github.io/}{hg-git}.}.
%
The use of DVCS in science will lead to long term productivity benefits
\citep{wilson_best_2014}.

@@

Note that open-source Matlab interpreters do exist, such as Octave and Scilab.
%
With some minimal effort, one can port existing Matlab codes using either
Julia which, by design, uses a Matlab-like syntax, or Python's \codeinline{pylab}
package provided via \pack{Matplotlib} and the command-line tool
\codeinline{fluidmat2py} in \pack{fluiddyn}.
%
For symbolic computation, some specialists use Maple or Mathematica, which are
good but expensive proprietary closed-source programs.
%
Sage and Sympy are two complementary open-source alternative both based on
Python.
%
The \fluidpack{lab} project in FluidDyn project is an attempt to orchestrate
experiments without relying on Labview.

@@

Some proper open-source codes have emerged, for example
\fnref{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
\fnref{https://www.openfoam.com/}{OpenFOAM} (C++),
\fnref{http://basilisk.fr}{Basilisk} (C) and
\fnref{http://channelflow.org/}{Channelflow} (C++) and
\fnref{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
%
Big companies have started to use open-source development for fluid mechanics
application. For example, Volkswagen group (Volkswagen, Audi, Seat, Porsche,
Skoda, ...) \fnref{%
  https://www.nas.nasa.gov/assets/pdf/ams/2014/AMS_20141202_Othmer.pdf}{%
  has used OpenFOAM since 2006}.
%
EDF, the main French electric utility company has also a interesting
\fnref{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
open-source strategy}. It has produced many programs and in particular the CFD
solver \fnref{https://www.code-saturne.org}{Code\_Saturn}.
%
\fnref{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}{%
EDF is also involved in the development of \fnref{https://www.paraview.org/}{Paraview}}.

Finally, scientists have started to use Python to study fluid mechanics,
especially through CFD codes (\fnref{http://dedalus-project.org/}{Dedalus},
\fnref{https://github.com/spectralDNS}{SpectralDNS},
\fnref{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
and \fnref{https://fenicsproject.org/}{FEniCS}) or through data analysis
(\fnref{http://www.openptv.net/}{OpenPTV},
\fnref{https://github.com/jr7/pypiv}{PyPIV}).
%
The packages of the FluidDyn project are also part of this trend.

>>>>

Collaboration was achieved in the early years of FOSS development through
emailing ``patches'', and centralized Version Control Systems (VCS) near the
turn of the 21\textsuperscript{st} century.  The collective collaboration on
development is today streamlined with the emergence of Distributed Version
Control Systems (DVCS), especially, Git and Mercurial).
%
Version control tools are designed to work on any kind of text files, not just
codes, so they could even be utilized to write scientific documents with
multiple authors.
%
In web-based source development platforms like GitHub, Bitbucket and GitLab,
DVCS repositories are enhanced with code-review tools, typically consisting of
a issue tracker and pull-request mechanism.
%
Git has become the \textit{de facto} standard for source management. However,
for the FluidDyn project, we prefer to use Mercurial because we consider it to
be simpler for the beginners and just as powerful for expert users\footnote{See
  {``a detailed comparison''} on Mercurial website,
  \url{https://www.mercurial-scm.org/wiki/GitConcepts}.
  }. Simplicity and low learning barriers for
scientists who are not experts in software development is crucial for our
project.
%
Note that, it is now possible to use Mercurial to work with Git repositories,
for example hosted in GitHub and GitLab\footnote{By using the Mercurial
extension \href{http://hg-git.github.io/}{hg-git}.}.
%
The use of DVCS in science will lead to long term productivity benefits
\citep{wilson_best_2014}.

@@

Note that open-source Matlab interpreters do exist, such as Octave and Scilab.
%
With some minimal effort, one can port existing Matlab codes using either
Julia which, by design, uses a Matlab-like syntax, or Python's \codeinline{pylab}
package provided via \pack{Matplotlib} and the command-line tool
\codeinline{fluidmat2py} in \pack{fluiddyn}.
%
For symbolic computation, some specialists use Maple or Mathematica, which are
good but expensive proprietary closed-source programs.
%
Sage and Sympy are two complementary open-source alternatives both based on
Python.
%
The \fluidpack{lab} project in FluidDyn project is an attempt to orchestrate
experiments without relying on Labview.

@@

Some proper open-source codes have emerged, for example
\fnref{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
\fnref{https://www.openfoam.com/}{OpenFOAM} (C++),
\fnref{http://basilisk.fr}{Basilisk} (C) and
\fnref{http://channelflow.org/}{Channelflow} (C++) and
\fnref{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
%
Big companies have started to use open-source development for fluid mechanics
application. For example, the automotive group Volkswagen \fnref{%
  https://www.nas.nasa.gov/assets/pdf/ams/2014/AMS_20141202_Othmer.pdf}{%
  has used OpenFOAM since 2006}.
%
EDF, the main French electric utility company has also a interesting
\fnref{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
open-source strategy}. It has produced many programs and in particular the CFD
solver \fnref{https://www.code-saturne.org}{Code\_Saturne}.
%
\fnref{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}{%
EDF is also involved in the development} of \fnref{https://www.paraview.org/}{Paraview}.

Finally, scientists have started to use Python to study fluid mechanics,
especially through CFD codes (\fnref{http://dedalus-project.org/}{Dedalus},
\fnref{https://github.com/spectralDNS}{SpectralDNS},
\fnref{https://github.com/mikaem/Oasis}{Oasis}, \fnref{http://pyfr.org/}{PyFR}
and \fnref{https://fenicsproject.org/}{FEniCS}) or through data analysis
(\fnref{http://www.openptv.net/}{OpenPTV},
\fnref{https://github.com/jr7/pypiv}{PyPIV}).
%
The packages of the FluidDyn project are also part of this trend.
